### 2017-01-31 Theory/Theory.tex
Some of my own editing...

\subsection{Fundamental particles}
\label{sec:theory_sm_particles}
In the \ac{SM} there are twelve fundamental fermions, six quarks and
six leptons. The quarks interact via both strong nuclear force and the 
electroweak force, while
the leptons only interact via the electoweak force.
Each of the fermions has a corresponding antiparticle
with opposite quantum numbers. The quarks and leptons
are divided into three generations. The particles from different
generations have similar properties, apart from the mass which increases
for each generation. The quarks and leptons are summarised in 
table \ref{tab:theory_fermions}.

@@

\subsection{Fundamental forces}
\label{sec:theory_sm_forces}
%Symmetries of the Lagrangian describing a physical system lead
%to conserved quantities, as shown by Noether's theorem \cite{noethersthm}.
%This is and important concept for the description of particle interactions
%under different forces: invariance (symmetry) under a certain 
%transformation group is imposed on 
%the Lagrangians describing the fundamental forces. This is brought
%about as local gauge invariance, where the Lagrangians 
%are required to be invariant under a local phase transformation
%plus a gauge transformation. The local phase transformation
%might introduce additional derivatives of potentials, taken care of by gauge transf.
The forces in the \ac{SM} are mediated by spin-1 bosons: The gluon (\Pgluon),
which carries the strong force, the photon (\Pphoton) as the
electromagnetic force carrier, and the $\PW^{\pm}$ and \PZ bosons, the
force carriers of the weak nuclear force.

The particles that interact via the strong force carry colour charge (red, green, or blue),
as do their mediators, the gluons. The interactions in the \ac{QFT} of the
strong force, called \ac{QCD}, are governed by an SU(3) (colour) symmetry.
%There are 8 gluons due to the SU(3) symmetry, colour octet and a colour singlet, but
%if there was a colour singlet it should occur as a free particle while it does not.
%See chapter 285 of griffiths
The quarks are never observed in isolation but always in bound
hadron states, which is referred to as confinement. Due to asymptotic
freedom \cite{asympt-I,asympt-II}, the strong coupling strength decreases at 
high energy (short distances). More detailed introductions to \ac{QCD} can 
be found in ref. \cite{griffiths}.

The weak and electromagnetic forces were unified into the
electroweak force. The unification of these forces was introduced 
in the 1960s by Glashow \cite{glashow-ewk}, Weinberg \cite{weinberg-ewk} and Salam \cite{salam-ewk},
is a key component of the \ac{SM}. Electroweak theory,
governed by an SU(2)$\times$U(1) symmetry, implies the presence of weak
neutral currents.

%Particles that carry electric charge interact via the
%electromagnetic force, while the weak force works
%between particles carrying weak isospin.
The weak interaction, working on all fermions,
only couples to the left-handed part of fermion states. In 
general a fermion state $\chi$ can be decomposed into its left- and right-handed parts
\begin{equation}\label{eqn:leftright}
\chi=\psi_R+\psi_L,
\end{equation}
and only $\psi_L$ will interact via the weak force. For massless particles
the general state $\chi$ is not a mixture of left-handed and right-handed parts,
but is either completely left-handed or right-handed. As such neutrinos, which are
massless in the \ac{SM}, must all be left-handed. %CAREFUL: handedness =/= helicity
The observation of neutrino flavour oscillations since the formulation of
the \ac{SM} implies that neutrinos do have a small but finite mass
and so there must be right-handed neutrinos in the \ac{SM}. This will
not be discussed here, for a more detailed account see reference \cite{pdg-2014}.
%The weak isospin acts on doublets of left-handed lepton pairs, while
%the right-handed leptons transform as singlet states. This works similarly
%for the quarks, with weak isospin acting three left-handed doublets and
Another feature of the weak force is that the weak eigenstates of the quarks
differ from the physical states, which mix via the \ac{CKM}-matrix \cite{ckm-matrix-cabibbo,ckm-matrix-KM}
in charged current interactions as
\begin{equation}\label{eqn:ckm_matrix}
\begin{pmatrix} d' \\
s' \\
b'  \end{pmatrix} = \begin{pmatrix} V_{ud} & V_{us} & V_{ub} \\
V_{cd} & V_{cs} & V_{cb} \\
V_{td} & V_{ts} & V_{tb} \end{pmatrix} \begin{pmatrix} d \\
s\\
b \end{pmatrix}.
\end{equation}

@@

Associated gauge fields are the three weak fields $W_{\mu}^{i}$ and the
weak hypercharge field $B_{\mu}$. These four fields mix to form the physical \Pphoton,
\PZ and $\PW^{\pm}$ states as
\begin{equation}\label{eqn:ewk_propagators_W}
\PW^{\pm} &= \frac{1}{\sqrt{2}}(W_{\mu}^1 \mp i W_{\mu}^2),
\end{equation}
\begin{equation}\label{eqn:ewk_propagators_phot}
A_{\mu} &= W_{\mu}^3\sin{\theta_W} + Y_{\mu}\cos{\theta_W},
\end{equation}
\begin{equation}\label{eqn:ewk_propagators_Z}
Z_{\mu} &= W_{\mu}^3\cos{\theta_W} - Y_{\mu}\sin{\theta_W}.
\end{equation}
In these equations $A_{\mu}$ is the photon field and $Z_{\mu}$ the \PZ field.
The weak mixing angle $\theta_W$ is related to the weak (g) and electromagnetic (g')
coupling constants as
\begin{equation}\label{eqn:thetaw}
\begin{split}
&g\sin{\theta_W} = g'\cos{\theta_W} \Rightarrow \\
&\sin{\theta_W} = \frac{g'}{\sqrt{g^2+g'^2}} \text{ and } \cos{\theta_W} = \frac{g}{\sqrt{g^2+g'^2}}
\end{split}
\end{equation}

@@

An issue with electroweak theory is that the gauge invariance of the Lagrangian would be
broken by the addition of mass terms for the gauge bosons. However, the \PW and \PZ bosons are known to be massive. 
Furthermore fermion mass terms would also not be gauge invariant, and it is also known
that the fermions are massive.
This means electroweak symmetry must be broken. In the \ac{SM}
this is achieved via the Higgs mechanism.

@@

The simplest addition to achieve this is a complex doublet,
\begin{equation}\label{eqn:phi_hfield}
\Phi = \begin{pmatrix} \phi^{+} \\
\phi^{0} \end{pmatrix}.
\end{equation}
This field has an associated Lagrangian
\begin{equation}\label{eqn:higgs_gen_lag}
\begin{split}
\mathcal{L}_{\text{Higgs}} = -(D_{\mu}\Phi)^{\dagger}(D^{\mu}\Phi) - V(\Phi^{\dagger}\Phi) \text{, where we choose, }\\
V = \lambda(\Phi^{\dagger}\Phi - \frac{\mu^2}{2\lambda})^2,
\end{split}
\end{equation}
Where $\lambda$ and $\mu^2$ are two real parameters, and $\lambda>0$ is needed for stability
of the vacuum. To get spontaneous symmetry breaking we require $\mu^2>0$.

Under a local $SU(2)\times U(1)$ transformation, with hypercharge Y=$\frac{1}{2}$
the resulting transformation of the field is,
\begin{equation}\label{eqn:hfield_trsf}
\begin{pmatrix} \phi^+ \\
\phi^0 \end{pmatrix} = \exp{\frac{i}{2}(\vec{\theta(x)}\cdot\vec{\sigma} + \rho(x))}
\begin{pmatrix} \phi^+ \\
\phi^0 \end{pmatrix}, 
\end{equation}
where $\vec{\sigma}$ are the Pauli spin matrices and $\vec{\theta(x)}$ and $\rho(x)$
are functions of the space-time coordinates. From such a transformation
the covariant derivative is required to be
\begin{equation}\label{eqn:H_cov_deriv}
D_{\mu} = \partial_{\mu} - \frac{i}{2}g\vec{W_{\mu}}\cdot\vec{\sigma} - \frac{i}{2}g'Y_{\mu}
\end{equation}
In the vacuum state we must have $V(\Phi^{\dagger}\Phi) =0$, which happens for nonzero values of $\Phi$,
and we can choose $\phi^+$ = 0 and $\phi^0$ real to get:
%\begin{equation}\label{eqn:vacuum_expt}
%\begin{split}
%&\Phi^{\dagger}\Phi - \frac{\mu^2}{2\lambda} = 0 \Rightarrow \\
%&\Phi^{\dagger}\Phi = \frac{\mu^2}{2\lambda} \Rightarrow\\
%&\phi^{+*}\phi^+ + \phi^{0*}\phi^0 = \frac{\mu^2}{2\lambda} \text{ and choosing } \phi^+ \text{ = 0 and } \phi^0 \text{ to be real,}\\
%&\phi^0 = \frac{v}{\sqrt{2}} \text{ where } v \text{ is real and} v^2=\frac{\mu^2}{\lambda}.
%\end{split}
%\end{equation}
\begin{equation}\label{eqn:field_vev}
\langle 0 | \Phi | 0 \rangle = \begin{pmatrix} 0 \\
\frac{v}{\sqrt{2}} \end{pmatrix},
\end{equation}
where $v$ is real and $v^2 = \frac{\mu^2}{\lambda}$.

Looking at fluctuations around the vacuum state,
\begin{equation}\label{eqn:fluctvac}
\begin{split}
&\delta\Phi = \frac{i}{2}(\vec{\theta}\cdot\vec{\sigma} + \rho)\Phi\\
&=\frac{i}{2}\begin{pmatrix} \theta_3 + \rho & \theta_1 - i\theta_2 \\
\theta_1+i\theta_2 & \rho - \theta_3 \end{pmatrix} \begin{pmatrix} 0\\
\frac{v}{\sqrt{2}} \end{pmatrix}.
\end{split}
\end{equation}
Requiring $\delta\Phi=0$, we have $\theta_1=\theta_2=0$ and $\rho=\theta_3$, 
and the only remaining generator that leaves the field 
invariant is $\frac{i}{2}\theta_3 + Y = I_3 +Y$ is the U(1) group. Denoting the 
remaining generators as $\vec{b}$, and parameterising the field as an expansion around the
vacuum state gives:
\begin{equation}\label{eqn:vev_expansion}
\Phi = \exp{\frac{i}{\sqrt{2}v}\vec{\theta}\cdot\vec{b}}\begin{pmatrix} 0 \\
\frac{1}{\sqrt{2}} (v+H) \end{pmatrix}.
\end{equation}
By choosing the appropriate gauge the Goldstone bosons $\vec{\theta}$
can be eliminated. They become the longitudinal degrees of freedom of the \PW and \PZ 
bosons. This leaves us with
\begin{equation}\label{eqn:vev_expansion_satisfied}
\Phi = \begin{pmatrix} 0 \\
\frac{1}{\sqrt{2}}(v+H) \end{pmatrix}
\end{equation}
Substituting this into the Lagrangian of equation \ref{eqn:higgs_gen_lag}, 
and using the definitions of $\PW_{\mu}^{\pm}$ and $\PZ_{\mu}$ from 
equations \ref{eqn:ewk_propagators}, we get
\begin{equation}\label{eqn:completed_lagrangian}
\mathcal{L}_{\text{Higgs}} = -\frac{1}{2}\partial_{\mu}H\partial^{\mu}H -\frac{g^2v^2}{4}\PW_{\mu}^- \PW^{\mu +} - \frac{g^2+g'^2}{8}v^2\PZ_{\mu}\PZ^{\mu}  - \mu^2H^2  + \text{terms of order 3 and up in the fields}.
\end{equation}
This gives us masses for the \PW bosons as $m_{\PW} = \frac{gv}{2}$ and a mass
for the \PZ boson as $m_{\PZ} = \frac{1}{2}v\sqrt{g^2+g'^2}$. The photon remains massles. 
The remaining mass term is for the field $H$, the
Higgs field, which has a mass $m_{H} = \sqrt{2\mu^2}$.
%Mass term : -1/2m_X^2. However for m_W note W_mu+Wmu- = 1/2(W_mu^1+iW_mu^2)(W_mu^1-iW_mu^2) = 
%1/2*(W_mu^1W_mu^1+W_mu^2W_mu^2). But also W+^2+W^-2 = W1^2+W2^2 so masses end up the same

@@

\section{Standard model Higgs boson measurements}
\label{sec:theory_smH}
The mass of the Higgs boson is a free parameter in the \ac{SM}, and so
searches for the boson had to be conducted over a large range of
masses. Searches for the Higgs boson at \ac{LEP} did not lead
to its observation. However, masses below 114.4 GeV were excluded at 95\% \ac{CL} \cite{LEP-Higgs},
providing a lower search limit. Subsequent studies at the 
Tevatron did also not lead to discovery of the Higgs boson, but did provide
an additional excluded masses range of 140-186 GeV at 95\% \ac{CL} \cite{TEV-Higgs}.

On the fourth of July 2012 the ATLAS and CMS collaborations
announced the discovery of a boson with a mass of around 125 GeV \cite{HDiscoveryATLAS,HDiscoveryCMS}.
The observations were based on excesses in the search for $\PHiggslight\rightarrow\gamma\gamma$ and $\PHiggslight\rightarrow \PZ\PZ$.
More data were collected and analysed during the remainder of 2012, and subsequent studies 
grew the confidence in this boson being compatible with the standard model Higgs boson. The ATLAS and CMS combined
mass measurement using the full dataset of Run 1 of the \ac{LHC} 
found \mh = $125.09 \pm 0.21 \text{ (stat) } \pm 0.11 \text{ (syst) }$ GeV \cite{MassComb}.
The combined ATLAS and CMS production- and decay rate measurement, also using the full dataset of \ac{LHC} Run 1,
shows very good agreement with the \ac{SM} predictions \cite{CouplComb}. In addition, through
this combination the $\PHiggslight \rightarrow \Pgt\Pgt$ decay was observed with a significance of 5.5$\sigma$, while
the individual experiments did not reach a significance upwards of 5$\sigma$ in this channel during Run 1.
Despite the good agreement with the \ac{SM}, some discrepancies do remain, 
and so the measurements of \ac{SM} Higgs boson
properties are certainly not finished yet.

@@

It is clear from the branching ratios of the \ac{SM} Higgs boson in figure \ref{fig:theory_smhxsbr}
that the two discovery channels $\PHiggslight \rightarrow \gamma\gamma$ and $\PHiggslight \rightarrow \PZ
PZ$ have smaller branching ratios than many other channels. These two channels are very sensitive
due to the small backgrounds and the excellent photon and lepton energy resolution the detectors provide.

During Run 2 of the LHC, standard model Higgs boson measurements, and searches for the 
standard model Higgs boson in decay channels not previously established, have 
already been performed. The decay in to $\gamma\gamma$ has been re-established \cite{CMSHgamgam2016,ATLASHgamgam2016},
as has the decay in to $ZZ\rightarrow 4\ell$ \cite{CMSHZZ2016,ATLASHZZ2016}. In addition, measurements
of $H\rightarrow WW$ are under way \cite{CMSHWW2016,ATLASHWW2016}. Establishment of the decay of 
$H\rightarrow b\bar{b}$ \cite{CMSVBFHbb2016,ATLASVHbb2016} and of $H\rightarrow \mu\mu$ \cite{ATLASHmm2016} 
is still to be achieved.
Finally, searches for ttH production, important to establish the top-Higgs coupling, have also been carried 
out \cite{CMSttH2016,CMSttHmultilep2016,ATLASttHbb2016,ATLASttHmultilep2016}. As the Run 2 dataset
keeps growing, more such measurements will be made to shed more light on the standard model Higgs boson.

\section{Beyond the standard model}
\label{sec:theory_BSM}
The \ac{SM} has been successfully tested to high accuraccy \cite{pdg-2014}. However,
some issues remain that cannot be addressed by the \ac{SM} alone \cite{griffiths}.
For example, it does not provide a candidate for dark matter. 
Additionally, the energy dependence of the running coupling constants
do not converge the GUT scale, and the force of gravity does not
appear in the \ac{SM}, no quantum theory can be formulated for it.

These issues aside, one of the most pressing issues regarding
the Higgs sector of the \ac{SM} is the hierarcy problem.
The \ac{SM}
is accepted to be an effective field theory that describes
the elementary particles and their interactions well up to an energy scale $\Lambda$.
This is the energy scale at which new physics must enter.
We know that $\Lambda \leq \mathcal{O}(10^{19})$, the Planck scale, as quantum 
gravitational effects start to become important in that region. $\Lambda$ enters
the corrections due to fermion and boson loops to the Higgs boson mass quadratically \cite{MSSM-carena-haber}:
\begin{equation}\label{eqn:mh_hierarchy}
m_{h_{\text{SM}}}^2  = (m_h^2)_0 + \mathcal{O}(\Lambda^2).
\end{equation}
Thus assuming that there is no new physics all the way up to
the Planck scale the observation \mh=125 GeV would only be compatible with $\Lambda$ of order $10^{19}$ GeV
if there was extreme fine-tuning of the bare Higgs boson mass $(m_h^2)_0$.

Many \acf{BSM} theories have been developed
that can to address the hierarchy problem and some of the other issues 
discussed. One of the most
popular of such theories is \acf{SUSY} \cite{SUSY-primer}, which postulates
that there is a symmetry between bosons and fermions. Every fermionic
\ac{SM} particle has a bosonic superpartner (named `s'+name of fermion, e.g.
sbottom, selectron,...), and every bosonic \ac{SM} particle
has a fermionic superpartner (named boson name + `ino', e.g. higgsino, Wino,...).

>>>>

\subsection{Fundamental particles}
\label{sec:theory_sm_particles}
In the \ac{SM} there are twelve fundamental fermions, six which
interact via the strong nuclear force (quarks) and six that do not (leptons).
Each of the fermions has a corresponding antiparticle
with opposite quantum numbers. The quarks and leptons
are divided into three generations, with corresponding particles from different
generations sharing the same properties apart from their mass.
The quarks and leptons are summarised in 
table \ref{tab:theory_fermions}.

@@

\subsection{Fundamental forces}
\label{sec:theory_sm_forces}
%Symmetries of the Lagrangian describing a physical system lead
%to conserved quantities, as shown by Noether's theorem \cite{noethersthm}.
%This is and important concept for the description of particle interactions
%under different forces: invariance (symmetry) under a certain 
%transformation group is imposed on 
%the Lagrangians describing the fundamental forces. This is brought
%about as local gauge invariance, where the Lagrangians 
%are required to be invariant under a local phase transformation
%plus a gauge transformation. The local phase transformation
%might introduce additional derivatives of potentials, taken care of by gauge transf.
The forces in the \ac{SM} are mediated by spin-1 bosons: the gluon (\Pgluon),
which carries the strong force, the photon (\Pphoton) 
mediating the electromagnetic force, and the $\PW^{\pm}$ and \PZ bosons, the
force carriers of the weak nuclear force.

The quantum number associated with the strong force is colour, which
is carried both by quarks and gluons.
The interactions in the \ac{QFT} of the
strong force, called \ac{QCD}, are governed by an SU(3) (colour) symmetry.
%There are 8 gluons due to the SU(3) symmetry, colour octet and a colour singlet, but
%if there was a colour singlet it should occur as a free particle while it does not.
%See chapter 285 of griffiths
Characteristics of \ac{QCD} include confinement, meaning quarks are always observed in bound
hadron states, and asymptotic
freedom \cite{asympt-I,asympt-II}. Due to asymptotic freedom the 
strong coupling strength decreases at 
high energy (or short distances), making the force weaker. 
A more detailed introduction to \ac{QCD} can 
be found in Ref. \cite{griffiths}.

The unification of the weak nuclear and electromagnetic forces was introduced 
in the 1960s by Glashow \cite{glashow-ewk}, Weinberg \cite{weinberg-ewk} and Salam \cite{salam-ewk}.
The electroweak theory is a key component of the \ac{SM}. It 
is governed by an SU(2)$\times$U(1) symmetry and implies the presence of weak
neutral currents.

%Particles that carry electric charge interact via the
%electromagnetic force, while the weak force works
%between particles carrying weak isospin.
%The weak force, which interacts with all fermions,
%only couples to the left-handed part of fermion states. In 
A general fermion state $\chi$ can be decomposed into its left- and right-handed parts
\begin{equation}\label{eqn:leftright}
\chi=\psi_R+\psi_L.
\end{equation}
The weak force only couples to $\psi_L$, and is thus violates parity maximally.
%where only $\psi_L$ will interact via the weak force. %For massless particles
%the general state $\chi$ is not a mixture of left-handed and right-handed parts,
%but is either completely left-handed or right-handed. As such neutrinos, which are
%massless in the \ac{SM}, must all be left-handed. %CAREFUL: handedness =/= helicity
%The observation of neutrino flavour oscillations since the formulation of
%the \ac{SM} implies that neutrinos do have a small but finite mass
%and so there must be right-handed neutrinos in the \ac{SM}. This will
%not be discussed here, for a more detailed account see reference \cite{pdg-2014}.
%The weak isospin acts on doublets of left-handed lepton pairs, while
%the right-handed leptons transform as singlet states. This works similarly
%for the quarks, with weak isospin acting three left-handed doublets and

Another feature of the weak force is that the weak eigenstates of the quarks
differ from the physical states, which mix via the \ac{CKM}-matrix \cite{ckm-matrix-cabibbo,ckm-matrix-KM}
in charged current interactions as
\begin{equation}\label{eqn:ckm_matrix}
\begin{pmatrix} d' \\
s' \\
b'  \end{pmatrix} = \begin{pmatrix} V_{ud} & V_{us} & V_{ub} \\
V_{cd} & V_{cs} & V_{cb} \\
V_{td} & V_{ts} & V_{tb} \end{pmatrix} \begin{pmatrix} d \\
s\\
b \end{pmatrix}.
\end{equation}

@@

Associated gauge fields are the three weak fields $W_{\mu}^{i}$ and the
weak hypercharge field $B_{\mu}$. These four fields mix to form the physical \Pphoton,
\PZ and $\PW^{\pm}$ states as
\begin{equation}\label{eqn:ewk_propagators_W}
\PW^{\pm} &= \frac{1}{\sqrt{2}}(W_{\mu}^1 \mp i W_{\mu}^2),
\end{equation}
\begin{equation}\label{eqn:ewk_propagators_phot}
A_{\mu} &= W_{\mu}^3\sin{\theta_W} + B_{\mu}\cos{\theta_W},
\end{equation}
\begin{equation}\label{eqn:ewk_propagators_Z}
Z_{\mu} &= W_{\mu}^3\cos{\theta_W} - B_{\mu}\sin{\theta_W}.
\end{equation}
In these equations $A_{\mu}$ is the photon field and $Z_{\mu}$ the \PZ field.
The weak mixing angle $\theta_W$ is related to the weak (g) and electromagnetic (g')
coupling constants as
\begin{equation}\label{eqn:thetaw}
\begin{split}
&g\sin{\theta_W} = g'\cos{\theta_W} \Rightarrow \\
&\sin{\theta_W} = \frac{g'}{\sqrt{g^2+g'^2}} \text{ and } \cos{\theta_W} = \frac{g}{\sqrt{g^2+g'^2}}
\end{split}
\end{equation}

@@

An issue with electroweak theory is that the gauge invariance of the Lagrangian would be
broken by the addition of mass terms for the gauge bosons. However, the \PW and \PZ bosons are known to be massive. 
A similar problem exists for fermions, as they are known to have mass but the introduction
of mass terms would break gauge invariance of the Lagrangian. 
This means electroweak symmetry must be broken. In the \ac{SM}
this is achieved via the Higgs mechanism.

@@

The simplest addition to achieve this is a complex doublet,
\begin{equation}\label{eqn:phi_hfield}
\Phi = \begin{pmatrix} \phi^{+} \\
\phi^{0} \end{pmatrix}.
\end{equation}
This field has an associated Lagrangian
\begin{equation}\label{eqn:higgs_gen_lag}
\begin{split}
&\mathcal{L}_{\text{Higgs}} = -(D_{\mu}\Phi)^{\dagger}(D^{\mu}\Phi) - V(\Phi^{\dagger}\Phi) \text{, where we choose, }\\
&V = \lambda(\Phi^{\dagger}\Phi - \frac{\mu^2}{2\lambda})^2,
\end{split}
\end{equation}
Where $\lambda$ and $\mu^2$ are two real parameters, and $\lambda>0$ is needed for stability
of the vacuum. To get spontaneous symmetry breaking we require $\mu^2>0$.

Under a local $SU(2)\times U(1)$ transformation, the field transforms as,
\begin{equation}\label{eqn:hfield_trsf}
\begin{pmatrix} \phi^+ \\
\phi^0 \end{pmatrix} = \exp{\frac{i}{2}(\vec{\theta}\cdot\vec{\sigma} + \rho)}
\begin{pmatrix} \phi^+ \\
\phi^0 \end{pmatrix}, 
\end{equation}
where $\vec{\sigma}$ are the Pauli spin matrices and $\vec{\theta}$ and $\rho$
are functions of the space-time coordinates. From such a transformation
the covariant derivative is required to be
\begin{equation}\label{eqn:H_cov_deriv}
D_{\mu} = \partial_{\mu} - \frac{i}{2}g\vec{W_{\mu}}\cdot\vec{\sigma} - \frac{i}{2}g'B_{\mu}
\end{equation}
In the vacuum state we must have $V(\Phi^{\dagger}\Phi) =0$, which happens for nonzero values of $\Phi$,
and we can choose $\phi^+$ = 0 and $\phi^0$ real to get:
%\begin{equation}\label{eqn:vacuum_expt}
%\begin{split}
%&\Phi^{\dagger}\Phi - \frac{\mu^2}{2\lambda} = 0 \Rightarrow \\
%&\Phi^{\dagger}\Phi = \frac{\mu^2}{2\lambda} \Rightarrow\\
%&\phi^{+*}\phi^+ + \phi^{0*}\phi^0 = \frac{\mu^2}{2\lambda} \text{ and choosing } \phi^+ \text{ = 0 and } \phi^0 \text{ to be real,}\\
%&\phi^0 = \frac{v}{\sqrt{2}} \text{ where } v \text{ is real and} v^2=\frac{\mu^2}{\lambda}.
%\end{split}
%\end{equation}
\begin{equation}\label{eqn:field_vev}
\langle 0 | \Phi | 0 \rangle = \begin{pmatrix} 0 \\
\frac{v}{\sqrt{2}} \end{pmatrix},
\end{equation}
where $v$ is real and $v^2 = \frac{\mu^2}{\lambda}$.

Considering infinitesimal fluctuations around the vacuum state, it can be shown that the
only remaining generator that leaves the field invariant is $I_3+Y$, the generator
of the U(1) group.
%\begin{equation}\label{eqn:fluctvac}
%\begin{split}
%&\delta\Phi = \frac{i}{2}(\vec{\theta}\cdot\vec{\sigma} + \rho)\Phi\\
%&=\frac{i}{2}\begin{pmatrix} \theta_3 + \rho & \theta_1 - i\theta_2 \\
%\theta_1+i\theta_2 & \rho - \theta_3 \end{pmatrix} \begin{pmatrix} 0\\
%\frac{v}{\sqrt{2}} \end{pmatrix}.
%\end{split}
%\end{equation}
%Requiring $\delta\Phi=0$, we have $\theta_1=\theta_2=0$ and $\rho=\theta_3$, 
%and the only generator that leaves the field invariant is $I_3+Y$, the generator of the U(1) group.
Denoting the 
remaining generators as $\vec{b}$, and parameterising the field as an expansion around the
vacuum state gives:
\begin{equation}\label{eqn:vev_expansion}
\Phi = \exp{\frac{i}{\sqrt{2}v}\vec{\theta}\cdot\vec{b}}\begin{pmatrix} 0 \\
\frac{1}{\sqrt{2}} (v+h) \end{pmatrix}.
\end{equation}
By choosing an appropriate gauge the Goldstone bosons $\vec{\theta}$
can be eliminated. They become the longitudinal degrees of freedom of the \PW and \PZ 
bosons. This leaves us with
\begin{equation}\label{eqn:vev_expansion_satisfied}
\Phi = \begin{pmatrix} 0 \\
\frac{1}{\sqrt{2}}(v+h) \end{pmatrix}
\end{equation}
Substituting this into the Lagrangian of equation \ref{eqn:higgs_gen_lag}, 
and using the definitions of $\PW_{\mu}^{\pm}$ and $\PZ_{\mu}$ from 
equations \ref{eqn:ewk_propagators_W}-\ref{eqn:ewk_propagators_Z}, we get
\begin{equation}\label{eqn:completed_lagrangian}
\begin{split}
&\mathcal{L}_{\text{Higgs}} = -\frac{1}{2}\partial_{\mu}h\partial^{\mu}h -\frac{g^2v^2}{4}\PW_{\mu}^- \PW^{\mu +} - \frac{g^2+g'^2}{8}v^2\PZ_{\mu}\PZ^{\mu}  - \mu^2h^2  + \\ 
&\text{terms of at least order 3 in the fields}.
\end{split}
\end{equation}
This leads to the \PW mass as $m_{\PW} = \frac{gv}{2}$ and a mass
for the \PZ boson as $m_{\PZ} = \frac{1}{2}v\sqrt{g^2+g'^2}$. Note that these
masses are related as $\frac{m_{\PW}}{m_{\PZ}} = \cos{\theta_W}$. The photon remains massles. 
The remaining mass term is for the field $h$, the
Higgs field, which has a mass $m_{h} = \sqrt{2\mu^2}$.
%Mass term : -1/2m_X^2. However for m_W note W_mu+Wmu- = 1/2(W_mu^1+iW_mu^2)(W_mu^1-iW_mu^2) = 
%1/2*(W_mu^1W_mu^1+W_mu^2W_mu^2). But also W+^2+W^-2 = W1^2+W2^2 so masses end up the same

@@

\section{Standard model Higgs boson measurements}
\label{sec:theory_smH}
The mass of the Higgs boson is a free parameter in the \ac{SM}, and so
a large range of possible masses had to be covered in searches. 
Searches for the Higgs boson at \ac{LEP} did not lead
to its observation. However, masses below 114.4 GeV were excluded at 95\% \ac{CL} \cite{LEP-Higgs},
providing a lower search limit. Subsequent studies at the 
Tevatron did also not lead to discovery of the Higgs boson, but did provide
an additional excluded masses range of 140-186 GeV at 95\% \ac{CL} \cite{TEV-Higgs}.

On the fourth of July 2012 the ATLAS and CMS collaborations
announced the discovery of a boson with a mass of around 125 GeV \cite{HDiscoveryATLAS,HDiscoveryCMS}.
The observations were based on excesses in the search for $\PHiggslight\rightarrow\gamma\gamma$ and $\PHiggslight\rightarrow \PZ\PZ$.
More data were collected and analysed during the remainder of 2012, and subsequent studies 
grew the confidence in the compatibility of this boson being the standard model Higgs boson. The ATLAS and CMS combined
mass measurement using the full dataset of Run 1 of the \ac{LHC} 
found its mass as \mh = $125.09 \pm 0.21 \text{ (stat) } \pm 0.11 \text{ (syst) }$ GeV \cite{MassComb}.
The combined ATLAS and CMS production- and decay rate measurement, also using the full dataset of \ac{LHC} Run 1,
shows very good agreement with the \ac{SM} predictions \cite{CouplComb}. In addition, through
this combination the $\PHiggslight \rightarrow \Pgt\Pgt$ decay was observed with a significance of 5.5$\sigma$, while
the individual experiments did not reach a significance upwards of 5$\sigma$ in this channel during Run 1.
%Despite the good agreement with the \ac{SM}, some discrepancies do remain, 
%and so the measurements of \ac{SM} Higgs boson
properties are certainly not finished yet.

@@

It is clear from the branching ratios of the \ac{SM} Higgs boson in figure \ref{fig:theory_smhxsbr}
that the two discovery channels $\PHiggslight \rightarrow \gamma\gamma$ and $\PHiggslight \rightarrow ZZ$
have smaller branching ratios than many other channels. These two channels are very sensitive
due to the small backgrounds and the excellent photon and lepton energy resolution the detectors provide.

During Run 2 of the LHC, standard model Higgs boson measurements, and searches for the 
standard model Higgs boson in decay channels not previously established, have 
already been performed. The decay in to $\gamma\gamma$ has been re-established \cite{CMSHgamgam2016,ATLASHgamgam2016},
as has the decay in to $ZZ\rightarrow 4\ell$ \cite{CMSHZZ2016,ATLASHZZ2016}, and measurements
of $h\rightarrow WW$ are being performed \cite{CMSHWW2016,ATLASHWW2016}. Establishment of the decay of 
$h\rightarrow b\bar{b}$ \cite{CMSVBFHbb2016,ATLASVHbb2016} and of $h\rightarrow \mu\mu$ \cite{ATLASHmm2016} 
is still to be achieved.
Finally, searches for ttH production, important to establish the top-Higgs coupling, have also been carried 
out \cite{CMSttH2016,CMSttHmultilep2016,ATLASttHbb2016,ATLASttHmultilep2016}. As the Run 2 dataset
keeps growing it will be possible to shed more light on the standard model Higgs boson.

\section{Beyond the standard model}
\label{sec:theory_BSM}
The \ac{SM} has been successfully tested to high accuraccy \cite{pdg-2014}. However,
some issues remain that cannot be addressed by the \ac{SM} alone \cite{griffiths}.
For example, it does not provide a candidate for dark matter.
In addition, the energy dependence of the running coupling constants
do not converge the GUT scale, and the force of gravity does not
appear in the \ac{SM}.

These points aside, the most important issue concerning
the Higgs sector of the \ac{SM} is the hierarchy problem.
The \ac{SM}
is accepted to be an effective field theory that describes
the elementary particles and their interactions well up to an energy scale $\Lambda$.
This is the energy scale at which new physics must enter.
We know that $\Lambda \leq \mathcal{O}(10^{19})$, the Planck scale, as quantum 
gravitational effects start to become important in that region. $\Lambda$ enters
the corrections due to fermion and boson loops to the Higgs boson mass quadratically \cite{MSSM-carena-haber}:
\begin{equation}\label{eqn:mh_hierarchy}
m_{h_{\text{SM}}}^2  = (m_h^2)_0 + \mathcal{O}(\Lambda^2).
\end{equation}
Thus assuming that there is no new physics all the way up to
the Planck scale the observation \mh=125 GeV would only be compatible with $\Lambda$ of order $10^{19}$ GeV
if there was extreme fine-tuning of the bare Higgs boson mass $(m_h^2)_0$.

Many \acf{BSM} theories that can address the hierarchy problem, and some of
the other issues that were mentioned, have been developed.
One of the most
popular of these theories is \acf{SUSY} \cite{SUSY-primer}, which postulates
that there is a symmetry between bosons and fermions. Every fermionic
\ac{SM} particle has a bosonic superpartner (sfermion, for scalar fermion)
and every bosonic \ac{SM} particle
has a fermionic superpartner (boson name + `ino', e.g. higgsino, Wino).

If this symmetry were unbroken the \ac{SUSY} particles must
have the same mass as their \ac{SM} partners. However, if that had 
been the case such particles would have been detected a long time ago. The 
only possibility is then that \ac{SUSY} is a broken symmetry and the superpartners
are heavier than their \ac{SM} equivalents. It is important to note
that the fermion and boson loops contribute to the Higgs boson
mass corrections with opposite sign, and so, if \ac{SUSY}
was an exact symmetry their contributions to the Higgs boson mass
would cancel. Due to 
symmetry breaking the contributions of \ac{SM} particles 
and their supersymmetric partners do not completely cancel. If
the \ac{SUSY} breaking scale, where new particles should
be found, is much larger than a few TeV further fine-tuning of 
the bare Higgs boson mass would be required \cite{MSSM-carena-haber,SUSY-primer}.
Apart from solving the hierarchy problem, \ac{SUSY} addresses some
of the other aforementioned issues. The lightest \ac{SUSY} particle
would be a candidate for dark matter if it were stable \cite{SUSY-primer}. On top of that,
\ac{SUSY} allows the electroweak and strong coupling
constants intersect at a common energy scale of $\mathcal{O}(10^{16} \text{ GeV })$ \cite{GUT-LEP}.
